---
title: "Practical Machine Learning Project"
author: Simon Hofer
date: 14.07.2015
output: html_document
---

## Prediction Assignment: Writeup


This is the course project for the Practical Machine Learning Course offered on Coursera by Johns Hopkins University. The following paragraph is copied directly from the course page.

### Information about Background and Data

#### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


#### Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 

### Loading the data and Cleaning

```{r}
library(caret)
set.seed(338833)
```

First, we load the data and clean it.
```{r}
# loading data and convert cells with no information to NA
training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))
# remove the first seven rows which do not contain relevent information
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
# delete rows with more than 1000 NA values in the training set
indices <- which(colSums(is.na(training)) < 1000)
training <- subset(training, select = indices)
testing <- subset(testing, select = indices)
```

### Machine Learning

After having cleaned the data and reduced the variables from 160 to 53, we now start with the machine learning.

Having a large data set at hand, we split it up in two subsets. This gives us the possibility to do a cross validation.

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
subtraining <- training[inTrain,]
subtesting <- training[-inTrain,]
```

Now, let's train the data with random forest.

*Important note: The model was made with the first bit of code in the following code chunk (marked as a comment). Because the machine learning took several hours, the file was saved locally and than important via the second bit of the code chunkt for performance issues.*

```{r}
# modFit <- train(classe ~ ., data=subtraining, method="rf", prox=T)
modFit <- readRDS("modFit.rds")
```

Finally, we predict our subtesting data set for cross validation.

```{r}
predsubtesting <- predict(modFit, subtesting)
confusionMatrix(predsubtesting, subtesting$classe)
```

The accuracy of our model is 99.7%, therefore the expected out-of-sample-error is 0.03%. The accuracy being really high, we expect that our model predicts the actual test data with none or very few errors.

## Prediction Assignment: Submission

First, we use our model to predict the classes of the 20 test instances. Than, we use the code from the course web site to write the 20 files, so we can upload them to the course server.

```{r}
predtesting <- predict(modFit, testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predtesting)
```